---
title: "Chen-Proposal"
output:
  pdf_document: default
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(infer)
library(dplyr)
library(readr)
```

================
Data Preparation
================
```{r}
# Load File
manydogs_2024 <- read_csv("~/Desktop/DATA606/manydogs_etal_2024_data.csv", show_col_types = FALSE)

# Set Seed
set.seed(1023)

# Clean and Prepare Data
# Filter out error and incomplete experiments
# Assign Yes/No to aggression answers to allow for mean calculations & aggregation
# Add up each row for aggression score
manydogs_2024_clean <- manydogs_2024 %>% filter(experiment_status=="Included") 
manydogs_2024_clean <- manydogs_2024_clean[c("environment", "owned_status", "sex", "age", "desexed", "purebred", "other_household_dogs", "num_household_dogs", "origin", "guardian_gender", "guardian_age", "cbarq_aggression_1","cbarq_aggression_2","cbarq_aggression_3","cbarq_aggression_4","cbarq_aggression_5","cbarq_aggression_6","cbarq_aggression_7","cbarq_aggression_8","cbarq_aggression_9","cbarq_aggression_10","cbarq_aggression_11","cbarq_aggression_12","cbarq_aggression_13","cbarq_aggression_14","cbarq_aggression_15","cbarq_aggression_16","cbarq_aggression_17","cbarq_aggression_18","cbarq_aggression_19","cbarq_aggression_20","cbarq_aggression_21","cbarq_aggression_22","cbarq_aggression_23","cbarq_aggression_24","cbarq_aggression_25","cbarq_aggression_26","cbarq_aggression_27")]


#create new agg_score using Yes/No methodology
for (i in 1:27) {
  var_name <- paste0("cbarq_aggression_", i)
  new_var  <- paste0("1agg", i)
  
  manydogs_2024_clean <- manydogs_2024_clean %>% 
    drop_na(all_of(var_name)) %>%  
    mutate(!!new_var := ifelse(.data[[var_name]] %in% c("No aggression", "Not observed"), 0, 1))
}

manydogs_2024_clean <- manydogs_2024_clean %>% mutate(agg_score = rowSums(across(starts_with("1agg")), na.rm = TRUE))


# Vector of aggression item names
agg_items <- paste0("cbarq_aggression_", 1:27)

# Lookup table
agg_map <- c(
  "No aggression"       = 0,
  "Not observed"        = 0,
  "Mild aggression"     = 1,
  "Moderate aggression" = 2,
  "Serious aggression"  = 3,
  "High aggression"     = 3
)

manydogs_2024_clean_v2 <- manydogs_2024_clean %>%
  dplyr::mutate(
    dplyr::across(
      .cols  = all_of(agg_items),
      .fns   = ~ unname(agg_map[.x]),
      .names = "alt_agg_{col}"     # <-- now names become: agg_cbarq_aggression_1
    )
  ) %>%
  dplyr::mutate(
    adjusted_agg_score = rowSums(dplyr::pick(starts_with("alt_agg_")), na.rm = TRUE)
  )

```



Revised Research question
What are factors which can lead to dog aggression.

What are the cases, and how many are there?
There are 704 cases (studies) collected from 20 research sites across eight countries.  After filtering for incomplete surveys/NAs, we are left with 155 observations which will be used.


Response: What is the response variable, and what type is it (numerical/categorical)?
There are 27 questions related to aggression which require the guardian to choose their response level: 
No aggression (No visible signs of aggression)
Mild aggression
Moderate aggression (Growling/barking/baring teeth)
High aggression
Serious aggression (Snaps, bites, or attempts to bite)
Not observed

===============================
Aggression score distributions:
===============================

The aggression score was computed two ways:
(1)Any scores moderate and above were counted as 1 and total was tallied
(2)Scores were ranked 0-3 with 3 being severe and total was tallied

The second option resulted in a wider disbursement of results & more outliers


```{r}
ggplot(data=manydogs_2024_clean_v2, aes(x=agg_score)) + geom_bar(fill="blue") + ggtitle(paste0("Distribution of Aggression Level"))
summary(manydogs_2024_clean_v2$agg_score)
ggplot(data=manydogs_2024_clean_v2, aes(x=adjusted_agg_score)) + geom_bar(fill="purple")+ ggtitle(paste0("Distribution of Aggression Level (adjusted)"))
summary(manydogs_2024_clean_v2$adjusted_agg_score)
```



********
Project Worksheet:

Section 1: Summary Stats

```{r}
#agg_score vs. adjusted_agg_score
ggplot(manydogs_2024_clean_v2, aes(x=adjusted_agg_score, y=agg_score)) + geom_point() + geom_smooth(se=FALSE)

ggplot(manydogs_2024_clean_v2, aes(x = environment, y = agg_score)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Aggression Scores by Environment",
       x = "Environment", y = "Aggression Score")

ggplot(manydogs_2024_clean_v2, aes(x = environment, y = adjusted_agg_score)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Aggression Scores by Environment (adjusted)",
       x = "Environment", y = "Adjusted Aggression Score")

```



Section 2: Regression Analysis

For adjusted agg score
```{r}
# full variables - adj R .1787
aggression_test <- lm(adjusted_agg_score ~ environment + owned_status + 
    sex + age + desexed + purebred + num_household_dogs + origin + 
    guardian_gender, data = manydogs_2024_clean_v2)
summary(aggression_test)

# remove environment, age, desexed,sex, guardian gender -  adj R .2034  -> Optimal for explanatory regression
aggression_test <- lm(adjusted_agg_score ~ owned_status + purebred + num_household_dogs + origin, data = manydogs_2024_clean_v2)
summary(aggression_test)

# added back age - adj .205
aggression_test <- lm(adjusted_agg_score ~ owned_status + purebred + age + num_household_dogs + origin, data = manydogs_2024_clean_v2)
summary(aggression_test)

# version 2 -adj R .1959
aggression_test_v2 <- lm(agg_score ~ + owned_status + purebred + age + num_household_dogs + origin, data = manydogs_2024_clean_v2)
summary(aggression_test_v2)
```

Regression analysis seems to indicate dogs from private homes with other dogs and the origin of the dog is from a shelter are best predictors of aggression.  However, the low regression score suggests these factors are limited in their ability to predict aggression behavior.  


Section 3: Regression Testing

```{r}
#residuals vs. fitted values
ggplot(data = aggression_test, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
  
#distribution of residuals
ggplot(data = aggression_test, aes(x = .resid)) +
  geom_histogram(bins = 25) +
  xlab("Residuals")

#normal prob plot of residuals  
ggplot(data = aggression_test, aes(sample = .resid)) +
  stat_qq()

ggplot(data = aggression_test, aes(x = owned_status, y = .resid)) +
  geom_boxplot(fill = "skyblue", color = "black")

ggplot(data = aggression_test, aes(x = num_household_dogs, y = .resid)) +
  geom_boxplot(fill = "skyblue", color = "black")

ggplot(data = aggression_test, aes(x = origin, y = .resid)) +
  geom_boxplot(fill = "skyblue", color = "black")

ggplot(data = aggression_test, aes(x = purebred, y = .resid)) +
  geom_boxplot(fill = "skyblue", color = "black")

ggplot(data = aggression_test, aes(x = age, y = .resid)) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)


ggplot(data=manydogs_2024_clean_v2, aes(x=owned_status, y=adjusted_agg_score)) + geom_boxplot(fill = "skyblue", color = "black")

ggplot(data=manydogs_2024_clean_v2, aes(x=num_household_dogs, y=adjusted_agg_score)) + geom_boxplot(fill = "skyblue", color = "black")

ggplot(data=manydogs_2024_clean_v2, aes(x=origin, y=adjusted_agg_score)) + geom_boxplot(fill = "skyblue", color = "black")

ggplot(data=manydogs_2024_clean_v2, aes(x=purebred, y=adjusted_agg_score)) + geom_boxplot(fill = "skyblue", color = "black")

ggplot(data=manydogs_2024_clean_v2, aes(x=age, y=adjusted_agg_score)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Section 4: ANOVA table

```{r}
# create ANOVA tables to test significance that the mean between groups is different.

#num_household_dogs  .0745*
model <- aov(adjusted_agg_score ~ num_household_dogs, data = manydogs_2024_clean_v2)
summary(model)

#environment  .707
model <- aov(adjusted_agg_score ~ environment, data = manydogs_2024_clean_v2)
summary(model)


#owned_status 2.59e-05**
model <- aov(adjusted_agg_score ~ owned_status, data = manydogs_2024_clean_v2)
summary(model)

#origin 0.001**
model <- aov(adjusted_agg_score ~ origin, data = manydogs_2024_clean_v2)
summary(model)

model <- aov(adjusted_agg_score ~ purebred, data = manydogs_2024_clean_v2)
summary(model)

#age 0.027*
model <- aov(adjusted_agg_score ~ age, data = manydogs_2024_clean_v2)
summary(model)

# envi vs. # household dogs .598
model <- aov(adjusted_agg_score ~ environment * num_household_dogs, data = manydogs_2024_clean_v2)
summary(model)

# owned status x number of dogs .593
model <- aov(adjusted_agg_score ~ owned_status * num_household_dogs, data = manydogs_2024_clean_v2)
summary(model)

```
Group means is slightly different
